---
title: "Modelling oak recruitment at abandoned croplands"
author: "Antonio J. Perez-Luque (@ajpelu)"
output:
  beamer_presentation:
    theme: "Pittsburgh"
    colortheme: "beaver"
    fig_height: 5
    fig_width: 7
    df_print: kable
    includes:
      in_header: header.tex
    incremental: no
    keep_tex: no
    latex_engine: xelatex
fontsize: 10pt
editor_options: 
  chunk_output_type: console
---


```{r packages, warning=FALSE, message=FALSE, echo=FALSE}
library(lattice)
library(tidyverse)
library(here)
library(knitr)
source(here::here("script/auxfx.R"))
## Text results
opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

## Code decoration
opts_chunk$set(tidy = FALSE, comment = NA, highlight = TRUE, size = "footnotesize")
```

```{r prepareData}
##  Read data
# Abandonded crop data
data_seed <- read.table(here::here("/datos/cul_abundance_seedling.txt"), head=TRUE, sep=";")

# Select only data of nseed.Qp  
ab.s <- data_seed[, c('nombre','localidad','tipo','subtipo','replicate','altitud','nseed.Qp')]

# Set 0 to NA 
ab.s[is.na(ab.s)] <- 0
```

## Aims 
Model Oak seedlings (_n.seed_; heigth < 130 cm) as a function of: 

- locality 
- type of ecosystem 
- Elevation 

## Data Exploration 

##### Outliers 

```{r DataExploration_A_OutliersY, echo=FALSE, fig.align='center', cache=TRUE}
# Y (seedling counts) (Cleveland dotplots)
par(mfrow = c(1, 2))
boxplot(ab.s$nseed.Qp, main='n.seed counts')
dotchart(ab.s$nseed.Qp, xlab='Range of data', ylab='n.seeds counts',
         groups=factor(ab.s$localidad), col=as.numeric(ab.s$tipo), pch=19)
# the groups bit ensures that observations from the same gruop are grouped together
# the pch command stands for point character (see pag 13. Zuur et al 2009)
```

## Data Exploration 

##### Outliers 
```{r DataExploration_A_OutliersX, echo=FALSE, fig.align='center', cache=TRUE}
# X variables 
MyVar <- c('altitud', 'localidad', 'tipo')
Mydotplot(as.matrix(ab.s[,MyVar]))  # See HighstatLibV6.R 
```

There are some potential outliers, in Canar locality. There are three transects with more than 50 seedling. 

```{r}
ab.s[ ab.s$nseed.Qp>50,]
```

These data are from forest type. 

## Data Exploration
##### Collinearity of covariables 

```{r DataExploration_B_Collinearity, fig.align='center', cache=TRUE}
## Collinearity 
MyVar <- c('altitud', 'localidad', 'tipo')
Mypairs(ab.s[, MyVar])
```

## Data Exploration: VIF

```{r echo=FALSE}
# Todas las covariables
corvif(ab.s[, MyVar])
# Quitamos la de mayor valor VIF
MyVar2 <- c('localidad','tipo')
corvif(ab.s[, MyVar2])
```

## Data Exploration
There is a high collinearity between ***altidud*** and ***localidad***. The variance inflation factors analysis reveals a medium value for ***altitud*** (vif=2.46). We removed the ***altitud*** from analysis (according to Zuur et al, 2013). We re-examined VIF and we found values close to 1 for the two variables. 

## Data Exploration
##### Relationships Y vs X

```{r DataExploration_C_Relationships, fig.align='center', cache=TRUE}
# If X is a factor....boxplots
par(mfrow = c(1, 2))
boxplot(nseed.Qp ~ localidad, data = ab.s, ylab='n seed counts', main='Relationship nseed.Qp ~ localidad')
abline(h=mean(ab.s$nseed.Qp),lty = 2)
boxplot(nseed.Qp ~ tipo, data = ab.s, ylab='n seed counts', main='Relationship nseed.Qp ~ tipo')
abline(h=mean(ab.s$nseed.Qp),lty = 2)
```

## Data Exploration
Another evidence of outliers for forest (`tipo='ROBLEDAL'`) of Canar locality (`localidad='CANAR'`).  

## Data Exploration
##### Spatial/temporal aspects of sampling design
not relevant here

## Data Exploration
##### Interactions 

```{r DataExploration_E_Interactions, fig.align='center', cache=TRUE}
# Interactions 
coplot(nseed.Qp ~ tipo | localidad, rows=1, data = ab.s)
```

```{r DataExploration_E_Interactions2, fig.align='center', cache=TRUE}
# Interactions 
bwplot(nseed.Qp ~ factor(tipo) | factor(localidad), data = ab.s, col = 1, layout=c(1,2))
```

## Data Exploration
##### Zero inflation Y 

```{r DataExploration_F_ZeroInflactions}
# F. Zero inflation
sum(ab.s$nseed.Qp == 0) / length(ab.s$nseed.Qp)
```

`r (sum(ab.s$nseed.Qp == 0) / length(ab.s$nseed.Qp))*100`% of data are zeros. We need to consider a model that accommodates this issue. 

## Data Exploration
##### Summary from Data Exploration

* There are three outliers. 
* The ***altitud*** variable has high collinearity and a elevated *vif*. 
* We decided to remove the three outliers and do not consider the altitud variable. 

## Data Exploration
 
```{r DataExploration_OUTLIER}
# Explore records with seelindg greather than 50
ab.s[ ab.s$nseed.Qp>50,]

# Remove data
ab.s1 <- ab.s[ab.s$nseed.Qp<50,]

# Explore again the interactions 
bwplot(nseed.Qp ~ factor(tipo) | factor(localidad), data = ab.s1, col = 1, layout=c(1,2))
```

